---
title: "rsnps 0.5.0: new features"
output: html_document
author: Sina Rüeger, Julia Gustavsen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
theme_set(theme_bw())
library(dplyr)
library(tidyr)
library(forcats)
```

## Summary

- `ncbi_snp_query()` now returns all reported allele frequencies in dbSNP (>= 0.5.0) in column `maf_population`.
- `ncbi_snp_query()` used to report only the allele frequency from [gnomAD](https://gnomad.broadinstitute.org/) (<= 0.4.0) in column `maf`.

## Changes

The [NEWS](https://github.com/ropensci/rsnps/blob/master/NEWS.md) file will tell you about the changes of rsnps 0.4.0 to 0.5.0, but the most relevant for users we would like to highlight in this blogpost. 

`ncbi_snp_query()` is the function that pulls data from [NCBI's](https://www.ncbi.nlm.nih.gov/) [dbSNP](https://www.ncbi.nlm.nih.gov/snp/), a database of single-nucleotide polymorphisms (SNP). This database lets a user query for a SNP of interest and returns a plethora of information, among them genomic position, associated gene, clinical significance and - relevant for this blogpost - the allele frequency. The allele frequency varies typically between different populations, sometimes just a little (e.g. [rs562556](https://www.ncbi.nlm.nih.gov/snp/rs562556#frequency_tab)), sometimes a lot (e.g. [rs11677783](https://www.ncbi.nlm.nih.gov/snp/rs11677783#frequency_tab)). This is why dbSNP collects allele frequency estimates from different studies and populations. Also, the database is constantly updated. 

Until version 0.4.0 `ncbi_snp_query()` reported the allele frequnecy estimated from gnomAD. For example, for SNP `rs420358` the `ncbi_snp_query` output used to look like this:

```r
ncbi_snp_query("rs420358")

#>          query chromosome        bp class         rsid          gene
#> 2     rs420358          1  40341239   snv     rs420358              

#>       alleles ancestral_allele variation_allele      seqname
#> 2     A,C,G,T                A            C,G,T NC_000001.11

#>     assembly ref_seq minor    maf
#> 2 GRCh38.p12       A     C 0.8614
```

We have now changed two things:

1. The `maf_population` column in the `ncbi_snp_query` output contains all reported allele frequencies, not only from one study or population (the `maf` column stays the same).
2. To do that, `ncbi_snp_query` returns now a tibble (and not a dataframe). 

## Examples

Let's have a look at the output of two SNPs mentioned above. You can see that `dat` is a) a tibble and b) that the column `maf_population` is a list.

```{r query}
# remotes::install_github("ropensci/rsnps", ref = "dev")
library(rsnps)
(dat <- ncbi_snp_query(c("rs11677783", "rs562556")))
```


Here is a visual to show how much the allele frequencies vary per SNP. 

First, we flatten the dataframe. 
```{r}
(dat_maf <- dat %>% select(query, maf_population) %>% unnest(cols = c(maf_population)))
```

Then we plot it. 
```{r plot_maf}


ggplot(data = dat_maf %>% filter(query == "rs11677783") %>% mutate(study = forcats::fct_reorder(study, MAF ))) + 
  geom_vline(xintercept = c(0, 0.5, 1), linetype = 3, color ="gray") +
  geom_point(aes(MAF, study)) + 
  labs(title = "Allele frequency", subtitle = "rs11677783") 

ggplot(data = dat_maf %>% filter(query == "rs562556") %>% mutate(study = forcats::fct_reorder(study, MAF ))) + 
  geom_vline(xintercept = c(0, 0.5, 1), linetype = 3, color ="gray") +
  geom_point(aes(MAF, study)) + 
  labs(title = "Allele frequency", subtitle = "rs562556") 

```

We can decide to turn the tibble into a dataframe again and pick a specific study (note that `maf_korean` is a `dbl` again): 

```{r}
(dat_korean <- 
  dat %>% mutate(maf_korean = purrr::map(maf_population, ~..1$MAF[..1$study=="KOREAN"])) %>% unnest(cols = c(maf_korean)))
                  
```

Or we can decide to pivot the study allele frequencies, so that each study has its own column:

```{r}
       
dat_maf <- dat %>% select(query, maf_population) %>% unnest(cols = c(maf_population)) %>%
  select(query, study, MAF) %>%
  pivot_wider(values_from = "MAF", names_from = "study", values_fn = min, names_prefix = "maf_") ## if duplicate, picking the minimum

(dat_wide <- 
  dat %>% left_join(dat_maf, by = "query"))


           
```
