---
output:
  pdf_document: default
  html_document: default
---
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rsnps tutorial}
%\VignetteEncoding{UTF-8}
-->

```{r eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# rsnps tutorial

## Install and load library

When available on CRAN

```{r install, eval=FALSE}
install.packages("rsnps")
```

Or get from Github

```{r installgithub, eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rsnps")
```


```{r load}
library(rsnps)
```

## OpenSNP data

### All Genotypes

Get genotype data for all users at a particular SNP from [OpenSNP](https://opensnp.org):

```{r allgensnp}
x <- allgensnp(snp='rs7412')
head(x)
```


### All Phenotypes

Get all phenotypes, their variations, and how many users have data available for a given phenotype

Get all data

```{r}
x <- allphenotypes(df = TRUE)
head(x)
```

Output a list, then call the characteristic of interest by 'id' or 'characteristic'

```{r}
datalist <- allphenotypes()
```

Get a list of all characteristics you can call

```{r}
names(datalist)[1:10]
```

Get data.frame for _ADHD_

```{r}
datalist[["ADHD"]]
```

Get data.frame for _mouth size_ and _SAT Writing_

```{r}
datalist[c("mouth size","SAT Writing")]
```

### Annotations

Get just the metadata

```{r annotations1}
annotations(snp = 'rs7903146', output = 'metadata')
```

Just from PLOS journals

```{r annotations2}
annotations(snp = 'rs7903146', output = 'plos')[c(1:2),]
```

Just from SNPedia

```{r annotations3}
annotations(snp = 'rs7903146', output = 'snpedia')
```

Get all annotations

```{r annotations4}
annotations(snp = 'rs7903146', output = 'all')[1:5,]
```

### Download

Download genotype data for a user from 23andme or other repo. (not evaluated in this example)

```{r fetch_genotypes, eval=FALSE}
data <- users(df=TRUE)
head(data[[1]])
fetch_genotypes(url = data[[1]][1,"genotypes.download_url"], rows=15)
```

### Genotype user data

Genotype data for one or multiple users

```{r}
genotypes(snp='rs9939609', userid=1)
```


```{r}
genotypes('rs9939609', userid='1,6,8', df=TRUE)
```


```{r}
genotypes('rs9939609', userid='1-2', df=FALSE)
```

### Phenotype user data

Get phenotype data for one or multiple users


```{r}
phenotypes(userid=1)$phenotypes[1:3]
```

```{r}
phenotypes(userid='1,6,8', df=TRUE)[[1]][1:10,]
```


```{r}
out <- phenotypes(userid='1-8', df=TRUE)
lapply(out, head)
```

### All known variations

Get all known variations and all users sharing that phenotype for one phenotype(-ID).

```{r}
phenotypes_byid(phenotypeid=12, return_ = 'desc')
```


```{r}
phenotypes_byid(phenotypeid=12, return_ = 'knownvars')
```


```{r}
phenotypes_byid(phenotypeid=12, return_ = 'users')[1:10,]
```


## NCBI SNP data

### dbSNP

Query NCBI's [dbSNP](https://www.ncbi.nlm.nih.gov/snp/) for information on a set of SNPs. 

An example with four markers, where one has been merged, and one has been withdrawn from NCBI.

```{r}
snps <- c("rs332", "rs420358", "rs1837253", "rs1209415715", "rs111068718")
(dbsnp_info <- ncbi_snp_query(snps))
```

The maf column contains the minor allele frequency from the GnomAD database (if available). All population specific allele frequencies can be accessed through the column `maf_population` which returns a list.
```{r}
dbsnp_info$maf_population
```