The `ropensnp` package interacts with the API services of [OpenSNP](http://opensnp.org/). No API key is needed. 

This tutorial will go through three use cases to demonstrate the kinds of things possible in `ropensnp`.

+ Get genotype data for all users at a particular snp.
+ Get genotype data for one or multiple users.
+ Get phenotype data for one or multiple users.
+ Get openSNP user information.

***
### Load package from GitHub
```{r loadpacakge}
# install_github("ropensnp", "ropensci")
library(ropensnp)
```

***

#### Get genotype data for all users at a particular snp.
```{r allgensnp}
head( allgensnp('rs7412', df=TRUE) )
```

***

#### Get genotype data for one or multiple users.
```{r genotypes}
genotypes(snp='rs9939609', userid=1)
genotypes('rs9939609', userid='1,6,8', df=TRUE)
```

***

### Get phenotype data for one or multiple users.
```{r phenotypes}
phenotypes(userid='1,6,8', df=TRUE)
head( phenotypes(userid='1-8', df=TRUE)[[1]] )

library(plyr)
head( ldply(phenotypes(userid='1-8', df=TRUE)) )
```


***

### Get openSNP user information.
```{r users}
# just the list
data <- users(df=FALSE)
data[1:2]

# get a data.frame of the users data
data <- users(df=TRUE)
head( data[[1]] ) # users with links to genome data
head( data[[2]] ) # users without links to genome data
```